---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

This data is found in the GitHub repository that was forked.

```{r}
activityfile <- read.csv("c:/Users/Vince/Documents/R/win-library/3.1/ReproducibleR/RepData_PeerAssessment1/activity/activity.csv")
```

To convert date into date format

```{r}
activityfile$date <- as.Date(activityfile$date)

```

## What is mean total number of steps taken per day?

* Ignore records with Steps = NA
```{r}
activityfile.narm <- activityfile[!is.na(activityfile$steps), ]
```

* Sum up number of steps by day
```{r}
stepsbyday <- rowsum(activityfile.narm$steps, activityfile.narm$date)
```

* Plot histogram
```{r}
hist(stepsbyday)
```

* Calculate mean
```{r}
mean(stepsbyday)
```

* Calculate median
```{r}
median(stepsbyday)
```


## What is the average daily activity pattern?

*Sum up number of steps by interval
```{r}
library(plyr)
stepsbyinterval <- ddply(activityfile.narm, ~interval, summarise, mean=mean(steps))
```

* generate plot
```{r}
plot(mean ~ interval, data=stepsbyinterval, type="l", main="Avg no of steps by interval")
```

*Find the interval with highest average no of steps

```{r}
stepsbyinterval[which.max(stepsbyinterval$mean), ]
```


## Imputing missing values

* Calculate total number of missing value in steps
```{r}
nrow(activityfile[is.na(activityfile$steps), ])
```

* Replace NA with mean in 5 min interval
```{r}
meanSteps <- aggregate(steps ~ interval, data = activityfile, FUN = mean)
imputeNA <- numeric()
for (i in 1:nrow(activityfile)) {
    row <- activityfile[i, ]
    if (is.na(row$steps)) {
        steps <- subset(meanSteps, interval == row$interval)$steps
    } else {
        steps <- row$steps
    }
    imputeNA <- c(imputeNA, steps)
}
```

* create new data
```{r}
activityfile.imputed <- activityfile
activityfile.imputed$steps <- imputeNA
```

* Create histogram of total no of steps taken each day
```{r}
totalSteps <- aggregate(steps ~ date, data = activityfile.imputed, sum, na.rm = TRUE)  
hist(totalSteps$steps, main = "Total steps by day")
```

* calculate mean
```{r}
mean(totalSteps$steps)
```

* calculate median
```{r}
median(totalSteps$steps)
```

## Are there differences in activity patterns between weekdays and weekends?

* create a variable indicating weekdays and weekend
```{r}
day <- weekdays(activityfile$date)
dayind <- vector()
for (i in 1:nrow(activityfile)) {
    if (day[i] == "Saturday") {
        dayind[i] <- "Weekend"
    } else if (day[i] == "Sunday") {
        dayind[i] <- "Weekend"
    } else {
        dayind[i] <- "Weekday"
    }
}
activityfile$dayind <- dayind
activityfile$dayind <- factor(activityfile$dayind)

stepsbyday <- aggregate(steps ~ interval + dayind, data = activityfile, mean)
names(stepsbyday) <- c("interval", "dayind", "steps")
```

* create a panel plot with time series
```{r}
library(lattice)
xyplot(steps ~ interval | dayind, stepsbyday, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")
```



